An Experimental Evaluation of Hybrid Vector Search (EA\&B)

### 2025-04-27 VLDB2026/main.tex
Update main.tex
table

	\begin{figure*}
		\centering
		\setlength{\abovecaptionskip}{0.1cm}
		\setlength{\belowcaptionskip}{-0.5cm}
		\includegraphics[width=0.92\textwidth]{figures/hybrid ANNS.pdf}
		\caption{Hybrid query example}

@@

	Despite BigANN 2023 \cite{bigann2023} evaluates the performance of some algorithms, it presents several notable limitations:
	1) The evaluation covers a limited number of algorithms and fails to comprehensively include mainstream methods. For instance, attribute filtering algorithms such as NHQ and UNG are not included in the evaluation. 2) It \textcolor{yellow}{focuses only on single-attribute and dual-attribute query scenarios,} neglecting the more complex demands of multi-attribute queries. In real-world applications, an object typically involves multiple attributes.\textcolor{yellow}{ For example, in e-commerce scenarios, users often specify multiple attributes (e.g., color, brand, price range) simultaneously when searching.}
	3) It does not evaluate algorithm performance under different attribute distributions. However, variations in attribute distribution may directly affect the filtering strategies and index construction efficiency of hybrid query algorithms.

@@

	\setlength{\textfloatsep}{0cm}
	\setlength{\floatsep}{0cm}
	\renewcommand{\arraystretch}{0.9}
	\begin{table*}[t]
		\centering
		%\setlength{\belowcaptionskip}{-0.3cm}
		\setlength{\abovecaptionskip}{0cm}
		\setlength{\belowcaptionskip}{-0.3cm}
		\caption{Comparison of AF-ANN search algorithms}
		\small
		\label{tab:compair_1}
		\begin{tabular}{|l|l|c|c|c|c|c|c|c|}
			\hline
			\textbf{Algorithm} & \textbf{Base} & \textbf{Filter Type} & \textbf{AND} & \textbf{OR} & \textbf{Flexible Attributes} & \textbf{Complex Boolean} & \textbf{Dynamic Insert} & \textbf{Multi Thread} \\
			\hline
			NHQ & Graph & C & Y & N& N& N & N & N \\
			FilteredVamana & Graph & C & N & Y & Y & N & Y & Y \\
			StitchedVamana & Graph & C & N & Y & Y & N & N & Y  \\
			CAPS & IVF & C & Y & N & Y & N & Y & Y \\
			ACORN & Graph & B & Y & Y & Y & Y & Y & Y \\
			UNG & Graph & B & Y & Y & Y & N & Y & Y \\ 
			Puck & IVF & B & Y & Y & Y & N & Y & Y \\
			% ParlayANN & Graph+IVF & B & N & N & Y & N & N & Y \\

			\hline

		\end{tabular}

		\centering
		%	 \setlength{\abovecaptionskip}{0cm}
		%	\setlength{\belowcaptionskip}{0.1cm}
		\footnotesize{
			\begin{minipage}{\linewidth}
				\vspace{0.1cm}
				\textsuperscript{A} Post-filtering, 
				\textsuperscript{B} Pre-filtering, 
				\textsuperscript{C} Simultaneous filtering, 
				\textsuperscript{Y} Support, 
				\textsuperscript{N} Unsupport. 
				\textit{Base} indicates that the hybrid query algorithm is an improvement based on a specific ANN search algorithm. 
				\textit{AND} indicates whether the algorithm supports attribute filtering with AND operations. 
				\textit{OR} indicates whether the algorithm supports attribute filtering with OR operations. 
				\textit{Flexible Attributes} indicates whether the number of query attributes can differ from the number of attributes used during index construction. 
				\textit{Complex Boolean} refers to whether the algorithm supports advanced Boolean logic beyond just AND and OR, such as NOT or combinations like (A AND B) OR (C AND NOT D). 
				\textit{Dynamic} indicates whether the algorithm supports the operation of dynamically inserting data points.
		\end{minipage}}
		\vspace{-0.5cm}
	\end{table*}

@@

	\setlength{\textfloatsep}{0.1cm}
	\setlength{\floatsep}{0cm}
		\renewcommand{\arraystretch}{0.9}
	\begin{table}[t]
		\centering
		\setlength{\abovecaptionskip}{0.1cm}
		\setlength{\belowcaptionskip}{-0.1cm}
		\setlength{\textfloatsep}{0.1cm}
		\caption{Comparison of RF-ANN search algorithms}
		\small	% 只会影响当前组的内容
		\label{tab:range_algo}
		\begin{tabular}{|c|c|c|c|}
			\hline
			\textbf{Algorithm} & \textbf{Base} & \textbf{Dynamic Insert} & \textbf{Multi Thread} \\
			\hline
			DSG & Graph & Y & N \\
			iRange & Graph & N & Y \\
			SeRF & Graph & Y & Y \\
			UNIFY & Graph & Y & Y \\
			WinFilter & Graph & N & Y  \\
			\hline
		\end{tabular}

		\footnotesize{
			\begin{minipage}{\linewidth}
				\textsuperscript{Y} Support, 
				\textsuperscript{N} Unsupport.
				\textit{Base} indicates that the hybrid query algorithm is an improvement based on a specific ANN search algorithm. 
				\textit{Dynamic} indicates whether the algorithm supports the operation of dynamically inserting data points.
			\end{minipage} 
			\vspace{-0.1cm}
		}

@@

	We introduce 5 representative range filtering algorithms and present a comparative overview in Table~\ref{tab:range_algo}.

@@

	\setlength{\textfloatsep}{0.1cm}
	\setlength{\floatsep}{0cm}
	\renewcommand{\arraystretch}{0.9}
	\begin{table}[t]
		\centering

		\setlength{\abovecaptionskip}{0cm}
		\setlength{\belowcaptionskip}{-0.3cm}
		\caption{Datasets}

		\label{tab:datasets_combined}
		\resizebox{\columnwidth}{!}{
			\begin{tabular}{cccccc}
				\toprule
				Dataset & Dimension & Base Data & Queries & LID & Type \\
				\midrule
				Msong & 420 & 992,272 & 200 & 23& Audio \\
				Audio & 192 & 53,387 & 200 & 14& Audio \\
				SIFT1M & 128 & 1,000,000 & 10,000 & 19& Image \\
				GIST1M & 960 & 1,000,000 & 1,000 & 45& Image \\
				GloVe & 100 & 1,183,514 & 10,000 & 47& Text \\
				Enron & 1369 & 94,987 & 200 & 23& Text \\
				Deep & 96 & 1,000,000 & 10,000 & 22& Image \\
				YT-Audio& 128 & 1,000,000 & 10,000 & 15& Audio \\
				WIT & 2048 & 1,000,000 & 40,300 & 48& Image \\
				Text2Image & 200 & 10,000,000 &10,000 & 59& Text2Image \\

@@

		\begin{subfigure}{\columnwidth}
			\centering
			\setlength{\abovecaptionskip}{0cm}
			\setlength{\belowcaptionskip}{-0.1cm}
			\includegraphics[width=0.95\linewidth]{figures/indexData/exp_7_build_time_comparison_query1.pdf}
			\caption{Index construction time}
			\label{fig:build_time_comparison_query1}
		\end{subfigure}

		\begin{subfigure}{\columnwidth}
			\centering
			\setlength{\abovecaptionskip}{0cm}
			\setlength{\belowcaptionskip}{-0.1cm}
			\includegraphics[width=0.95\linewidth]{figures/indexData/exp_7_index_size_mb_comparison_query1.pdf}
			\caption{Index size}
			\label{fig:index_size_mb_comparison_query1}
		\end{subfigure}

		\begin{subfigure}{\columnwidth}
			\centering
			\setlength{\abovecaptionskip}{0cm}
			\setlength{\belowcaptionskip}{-0.3cm}
			\includegraphics[width=0.95\linewidth]{figures/indexData/exp_7_memory_mb_comparison_query1.pdf}
			\caption{Peak memory usage}
			\label{fig:memory_mb_comparison_query1}
		\end{subfigure}

@@

	\begin{figure*}
		\centering
		        \setlength{\abovecaptionskip}{0cm}
		    \setlength{\belowcaptionskip}{-0.5cm}
		\includegraphics[width=0.98\textwidth]{figures/exp/exp_1_1_SingleLabel_1thread.pdf}
		\caption{Single-Attribute Building and Query (single thread) }
		\label{fig:exp_1_1_SingleLabel_1thread}
	\end{figure*}

	\begin{figure*}
		\centering
		    \setlength{\abovecaptionskip}{0cm}
		    \setlength{\belowcaptionskip}{-0.5cm}
		\includegraphics[width=0.98\textwidth]{figures/exp/exp_1_2_SingleLabel_16thread.pdf}
		\caption{Single-Attribute Building and Query (16 threads)}
		\label{fig:exp_1_2_SingleLabel_16thread}
	\end{figure*}

@@

	\begin{figure}[t]
		\centering
		\setlength{\abovecaptionskip}{0.1cm}
		\setlength{\belowcaptionskip}{-0.1cm}
		% 上面的图例，居中并可通过 hspace 调整左右位置
		\hspace*{15pt} % 可调整的参数，负值左移，正值右移，例如 -20pt 或 20pt
		\includegraphics[width=0.95\columnwidth]{figures/exp/exp_2_legend.pdf} % 图例图片路径

		% 下面的子图，居中显示
		\includegraphics[width=0.7\columnwidth]{figures/exp/exp_2_1.pdf}
		\caption{Effect of Multi-Attribute Index Construction on Single Query Performance. ($S$ denotes single-attribute index construction and single-attribute query; $M$ denotes multi-attribute index construction and single-attribute query)}
		\label{fig:exp_2_1}

@@

				\hline
				S5: General-purpose Use & Milvus, Faiss &DSG, UNIFY\\
				\hline
				S6: Resource-constrained Environments & CAPS, Faiss&Faiss, SeRF \\
				\hline
				S7: High-Recall Querying & UNG & WinFilter \\
				\hline
				S8: OOD scenarios & UNG &WinFilter  \\
				\hline
			\end{tabular}
		}
		\label{tab:algo_scenarios}

>>>>

	\begin{figure*}
		\centering
		\setlength{\abovecaptionskip}{0.1cm}
		\setlength{\belowcaptionskip}{-0.5cm}
		\includegraphics[width=\textwidth]{figures/hybrid ANNS.pdf}
		\caption{Hybrid query example}

@@

	Despite BigANN 2023 \cite{bigann2023} evaluates the performance of some algorithms, it presents several notable limitations:
	1) The evaluation covers a limited number of algorithms and fails to comprehensively include mainstream methods. For instance, attribute filtering algorithms such as NHQ and UNG are not included in the evaluation. 2) It focuses only on single-attribute and dual-attribute query scenarios, neglecting the more complex demands of multi-attribute queries. In real-world applications, an object typically involves multiple attributes. For example, in e-commerce scenarios, users often specify multiple attributes (e.g., color, brand, price range) simultaneously when searching.
	3) It does not evaluate algorithm performance under different attribute distributions. However, variations in attribute distribution may directly affect the filtering strategies and index construction efficiency of hybrid query algorithms.

@@

	\setlength{\textfloatsep}{0cm}
	\setlength{\floatsep}{0cm}
	\begin{table*}[t]
		\centering
		%\setlength{\belowcaptionskip}{-0.3cm}
		\setlength{\abovecaptionskip}{0.1cm}
		\setlength{\belowcaptionskip}{-0.2cm}
		\caption{Comparison of AF-ANN search algorithms}
		\small
		\label{tab:compair_1}
%		\begin{tabular}{|l|l|c|c|c|c|c|c|c|}
%			\hline
%			\textbf{Algorithm} & \textbf{Base} & \textbf{Filter Type} & \textbf{AND} & \textbf{OR} & \textbf{Flexible Attributes} & \textbf{Complex Boolean} & \textbf{Dynamic Insert} & \textbf{Multi Thread} \\
%			\hline
%			NHQ & Graph & C & Y & N& N& N & N & N \\
%			FilteredVamana & Graph & C & N & Y & Y & N & Y & Y \\
%			StitchedVamana & Graph & C & N & Y & Y & N & N & Y  \\
%			CAPS & IVF & C & Y & N & Y & N & Y & Y \\
%			ACORN & Graph & B & Y & Y & Y & Y & Y & Y \\
%			UNG & Graph & B & Y & Y & Y & N & Y & Y \\ 
%			Puck & IVF & B & Y & Y & Y & N & Y & Y \\
%			% ParlayANN & Graph+IVF & B & N & N & Y & N & N & Y \\
%			
%			\hline
%			
%		\end{tabular}
	\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|l|l|}
		\hline
		\textbf{Item} & \textbf{NHQ} & \textbf{FilteredVamana} & \textbf{StitchedVamana} & \textbf{CAPS} & \textbf{ACORN} & \textbf{UNG} & \textbf{Puck} & \textbf{DSG} & \textbf{iRange} & \textbf{SeRF} & \textbf{UNIFY} & \textbf{WinFilter} \\
		\hline
		Base & Graph & Graph & Graph & IVF & Graph & Graph & IVF & Graph & Graph & Graph & Graph & Graph \\
		Dynamic Insert & N & Y & N & Y & Y & Y & Y & Y & N & Y & Y & N \\
		Multi Thread & N & Y & Y & Y & Y & Y & Y & N & Y & Y & Y & Y \\
		Filter Type & C & C & C & C & B & B & B & - & - & - & - & - \\
		AND & Y & N & N & Y & Y & Y & Y & - & - & - & - & - \\
		OR & N & Y & Y & N & Y & Y & Y & - & - & - & - & - \\
		Flexible Attributes & N & Y & Y & Y & Y & Y & Y & - & - & - & - & - \\
		Complex Boolean & N & N & N & N & Y & N & N & - & - & - & - & - \\

		\hline
	\end{tabular}

		\centering
		%	 \setlength{\abovecaptionskip}{0cm}
		%	\setlength{\belowcaptionskip}{0.1cm}
		\footnotesize{
			\begin{minipage}{\linewidth}
				\vspace{0.1cm}
				\textsuperscript{A} Post-filtering, 
				\textsuperscript{B} Pre-filtering, 
				\textsuperscript{C} Simultaneous filtering, 
				\textsuperscript{Y} Support, 
				\textsuperscript{N} Unsupport. 
				\textit{Base} indicates that the hybrid query algorithm is an improvement based on a specific ANN search algorithm. 
				\textit{AND} indicates whether the algorithm supports attribute filtering with AND operations. 
				\textit{OR} indicates whether the algorithm supports attribute filtering with OR operations. 
				\textit{Flexible Attributes} indicates whether the number of query attributes can differ from the number of attributes used during index construction. 
				\textit{Complex Boolean} refers to whether the algorithm supports advanced Boolean logic beyond just AND and OR, such as NOT or combinations like (A AND B) OR (C AND NOT D). 
				\textit{Dynamic} indicates whether the algorithm supports the operation of dynamically inserting data points.
		\end{minipage}}
		\vspace{-0.3cm}
	\end{table*}

@@

%	\setlength{\textfloatsep}{0.1cm}
%	\setlength{\floatsep}{0cm}
%	\begin{table}[t]
%		\centering
%		\setlength{\abovecaptionskip}{0.1cm}
%		\setlength{\belowcaptionskip}{-0.1cm}
%		\setlength{\textfloatsep}{0.1cm}
%		\caption{Comparison of RF-ANN search algorithms}
%		\small	% 只会影响当前组的内容
%		\label{tab:range_algo}
%		\begin{tabular}{|c|c|c|c|}
%			\hline
%			\textbf{Algorithm} & \textbf{Base} & \textbf{Dynamic Insert} & \textbf{Multi Thread} \\
%			\hline
%			DSG & Graph & Y & N \\
%			iRange & Graph & N & Y \\
%			SeRF & Graph & Y & Y \\
%			UNIFY & Graph & Y & Y \\
%			WinFilter & Graph & N & Y  \\
%			\hline
%		\end{tabular}
%		
%		
%		\footnotesize{
%			\begin{minipage}{\linewidth}
%				\textsuperscript{Y} Support, 
%				\textsuperscript{N} Unsupport.
%				\textit{Base} indicates that the hybrid query algorithm is an improvement based on a specific ANN search algorithm. 
%				\textit{Dynamic} indicates whether the algorithm supports the operation of dynamically inserting data points.
%			\end{minipage} 
%		}
%		
%	\end{table}

@@

	We introduce 5 representative range filtering algorithms and present a comparative overview in Table~\ref{tab:compair_1}.

@@

	\setlength{\textfloatsep}{0.1cm}
	\setlength{\floatsep}{0cm}
	\begin{table}[t]
		\centering

		\setlength{\abovecaptionskip}{0.05cm}
		\setlength{\textfloatsep}{0.cm}
		\caption{Datasets}

		\label{tab:datasets_combined}
		\resizebox{\columnwidth}{!}{
			\begin{tabular}{cccccc}
				\toprule
				Dataset & Dimension & Base Data & Queries & LID & Type \\
				\midrule
				Msong & 420 & 992,272 & 200 & 23& Audio \\[2pt]
				Audio & 192 & 53,387 & 200 & 14& Audio \\[2pt]
				SIFT1M & 128 & 1,000,000 & 10,000 & 19& Image \\[2pt]
				GIST1M & 960 & 1,000,000 & 1,000 & 45& Image \\[2pt]
				GloVe & 100 & 1,183,514 & 10,000 & 47& Text \\[2pt]
				Enron & 1369 & 94,987 & 200 & 23& Text \\[2pt]
				Deep & 96 & 1,000,000 & 10,000 & 22& Image \\[2pt]
				YT-Audio& 128 & 1,000,000 & 10,000 & 15& Audio \\[2pt] 
				WIT & 2048 & 1,000,000 & 40,300 & 48& Image \\[2pt]
				Text2Image & 200 & 10,000,000 &10,000 & 59& Text2Image \\[2pt]

@@

		\begin{subfigure}{\columnwidth}
			\centering
			\setlength{\abovecaptionskip}{0cm}
			\setlength{\belowcaptionskip}{-0.1cm}
			\includegraphics[width=\linewidth]{figures/indexData/exp_7_build_time_comparison_query1.pdf}
			\caption{Index construction time}
			\label{fig:build_time_comparison_query1}
		\end{subfigure}

		\begin{subfigure}{\columnwidth}
			\centering
			\setlength{\abovecaptionskip}{0cm}
			\setlength{\belowcaptionskip}{-0.1cm}
			\includegraphics[width=\linewidth]{figures/indexData/exp_7_index_size_mb_comparison_query1.pdf}
			\caption{Index size}
			\label{fig:index_size_mb_comparison_query1}
		\end{subfigure}

		\begin{subfigure}{\columnwidth}
			\centering
			\setlength{\abovecaptionskip}{0cm}
			\setlength{\belowcaptionskip}{-0.3cm}
			\includegraphics[width=\linewidth]{figures/indexData/exp_7_memory_mb_comparison_query1.pdf}
			\caption{Peak memory usage}
			\label{fig:memory_mb_comparison_query1}
		\end{subfigure}

@@

	\begin{figure*}
		\centering
		        \setlength{\abovecaptionskip}{0cm}
		    \setlength{\belowcaptionskip}{-0.5cm}
		\includegraphics[width=\textwidth]{figures/exp/exp_1_1_SingleLabel_1thread.pdf}
		\caption{Single-Attribute Building and Query (single thread) }
		\label{fig:exp_1_1_SingleLabel_1thread}
	\end{figure*}

	\begin{figure*}
		\centering
		    \setlength{\abovecaptionskip}{0cm}
		    \setlength{\belowcaptionskip}{-0.5cm}
		\includegraphics[width=\textwidth]{figures/exp/exp_1_2_SingleLabel_16thread.pdf}
		\caption{Single-Attribute Building and Query (16 threads)}
		\label{fig:exp_1_2_SingleLabel_16thread}
	\end{figure*}

@@

	\begin{figure}[th]
		\centering
		\setlength{\abovecaptionskip}{0.1cm}
		\setlength{\belowcaptionskip}{-0.1cm}
		% 上面的图例，居中并可通过 hspace 调整左右位置
		\hspace*{15pt} % 可调整的参数，负值左移，正值右移，例如 -20pt 或 20pt
		\includegraphics[width=0.98\columnwidth]{figures/exp/exp_2_legend.pdf} % 图例图片路径

		% 下面的子图，居中显示
		\includegraphics[width=0.8\columnwidth]{figures/exp/exp_2_1.pdf}
		\caption{Effect of Multi-Attribute Index Construction on Single Query Performance. ($S$ denotes single-attribute index construction and single-attribute query; $M$ denotes multi-attribute index construction and single-attribute query)}
		\label{fig:exp_2_1}

@@

				\hline
				S5: General-purpose Use & Milvus, Faiss &DSG, UNIFY\\
				\hline
				S6: Resource constraint & CAPS, Faiss&Faiss, SeRF \\
				\hline
				S7: High-Recall Querying & UNG & WinFilter \\
				\hline
				S8: OOD scenarios & UNG &WinFilter  \\
				\hline
			\end{tabular}
		}
		\label{tab:algo_scenarios}
